Modelling Stochastic Time Delay for Regression

### 2021-04-20 documentation/article/article.tex

\title{Modelling Stochastic Time Delay for Regression Analysis}
\author[Aaron Pickering]{Aaron Pickering}
\email{\href{mailto:aaron1rcl@gmail.com}{aaron1rcl@gmail.com}}
\author[Juan Camilo Orduz]{Juan Camilo Orduz}
\email{\href{mailto:juanitorduz@gmail.com}{juanitorduz@gmail.com}}
\urladdr{\href{https://juanitorduz.github.io/}{juanitorduz.github.io}}
%\address{}

@@

\begin{center}
\begin{figure}
\includegraphics[totalheight=6cm]{images/2_input_delay_example.png}
\caption{Add figure caption}
\label{fig:verticalcell}
\end{figure}
\end{center}

@@

First, let us iniialise some starting values for each of these parameters. Now, we want to find the best possible time shift $\tau$ for each input impulse in $X(t)$. It stands to reason that the best possible time shift would be one that is not too distant from the observed impulse (i.e. has a high likelihood given some distribution) and also produces the best possible prediction. From there, the likelihood estimate is derived from the time shift and the prediction error.
Finally, we iterate over a number of values of $\tau$ optimising until we maximise the likelihood for the specific impulse.

@@

Accordingly, the optimisation procedure is (CHANGE TO GENERAL PARAMETERS):

\begin{enumerate}
    \item [(1)] Initialise the set of parameters $\beta$, $\sigma_{\varepsilon}$ and $\lambda_{\tau}$ for the function $f$.
    \item [(2)] Find the $\mathcal{T}$ which maximises the likelihood for the given set of parameters:
        \begin{enumerate} 
            \item[(i)] Initialize the discrete optimisation algorithm with all parameters set to the mean of the $\mathcal{T}$ distribution ($\lambda_{\tau}$).
            \item[(ii)] Next, randomly select a small number of impulses $m$, with the value of $m$ treated as a hyperparameter. For each impulse, a random time shift is drawn from the $\mathcal{T}$ distribution creating a proposal vector. 
            \item [(iii)]The likelihood (both time shift and prediction error) for the proposal is calculated. 
            \item[(iv)] If the proposal likelihood is higher than the current maximum likelihood, our estimate is updated.
            \item [(v)] Return to (i) and repeat. The best estimate of the set of $\mathcal{T}$ improves each iteration. The number of iterations ($N$) is also treated as a hyperparameter.
        \end{enumerate} 
    \item [(3)] Optimise the model parameters $\beta$, $\sigma_{\varepsilon}$ and $\lambda_{\tau}$.
\end{enumerate}

@@

\subsection{Constraints (TO DO: ReWrite)}
Finally, a note on problem constraints. The likelihood estimates are dependent on the inner optimisation procedure. There is no guarantee that the global maximum will be found, particularly for sequences with a high density of impulses. For example, consider the sequence:

\begin{align*}
    x(t) = 
    \left(
\begin{array}{cccccccccc}
    0.5 & 0.6 & 0.4 & 1 & 2 & 1 &
    0.3 & 0.2 & 0.5 & 0.8 
\end{array}  
\right)
\end{align*}

where each impulse can shift between -2 and +2 positions. Ignoring shifts beyond the edge of the vector, the solutions space of $\mathcal{T}$ is $5^t$ combinations.  
In such cases, the estimated likelihood is likely to be close to but not exactly the same as the real value. As the density (in time) of impulses increases, the accuracy of the estimates decrease. Therefore, this model is most appropriate for sparse time series inputs.

@@

The following section demonstrates a simulated univariate example of TVS Regression. Figure \ref{fig:simulated_example} shows the simulated time series. The full code can be found via \href{https://github.com/aaron1rcl/tvs_regression/blob/master/notebooks/1_univariate_example.ipynb}{TVS Regression on GitHub}. 
The input signal has $20$ non-zero impulses which have been drawn from the standard normal distribution. The system is 'off' when $x(t) =0$. In other words, when $x(t) = 0$, $f(x(t)) = 0$. The true shift distribution $\mathcal{T}$ is given by $\mathcal{T} \sim Poisson(\lambda = 2)$. There is one value of $\tau$ for each of the 20 impulses. The green line in Figure \ref{fig:simulated_example} represents the shifted series, corresponding to the time at which the effect occurs. The blue line is the output $y$ which includes a small amount of gaussian noise. The values for the parameters were arbitrarily selected. The output $y$ is defined by the following equation:

@@

\begin{center}
\begin{table}
\begin{tabular}{|c|c|c|c|}
\hline
Parameter & TVS Regression & OLS Regression & True Value \\
\hline
$\beta$ & 2.09 & 0.50 & 2.00  \\
\hline
$Intercept$ & 6.53 & 6.62 & 6.50 \\
\hline
$\lambda_{\tau}$ & 1.54 & $N/A$ & 1.40 \\
\hline
$\sigma_{\epsilon}$ & 0.20 & 1.03 & 0.20 \\
\hline
\end{tabular}
\caption{ Parameter comparison between TVR and OLS regression against the true value.} 
\label{results_table}
\end{table}
\end{center}

>>>>

\title{Modelling Stochastic Time Delay for Regression}
\author[Aaron Pickering]{Aaron Pickering}
\email{\href{mailto:aaron_rcl@hotmail.com}{aaron\_rcl@hotmail.com}}
\author[Juan Camilo Orduz]{Juan Camilo Orduz}
\email{\href{mailto:juanitorduz@gmail.com}{juanitorduz@gmail.com}}
\urladdr{\href{https://juanitorduz.github.io/}{juanitorduz.github.io}}
%\address{}

@@

\begin{center}
\begin{figure}
\includegraphics[totalheight=6cm]{images/2_input_delay_example.png}
\caption{Stochastic Time Delay Example}
\label{fig:verticalcell}
\end{figure}
\end{center}

@@

First, let us initialise some starting values for each of these parameters. Now, we want to find the best possible time shift $\tau$ for each input impulse in $X(t)$. It stands to reason that the best possible time shift would be one that is not too distant from the observed impulse (i.e. has a high likelihood given some distribution) and also produces the best possible prediction. From there, the likelihood estimate is derived from the time shift and the prediction error.
Finally, we iterate over a number of values of $\tau$ optimising until we maximise the likelihood for the specific impulse.

@@

Accordingly, the optimisation procedure is:

\begin{enumerate}
    \item [(1)] Initialise the set of parameters $\theta_{f}$ and $\theta_{\mathcal{T}}$ for the function $f$.
    \item [(2)] Find the $\mathcal{T}$ which maximises the likelihood for the given set of parameters:
        \begin{enumerate} 
            \item[(i)] Initialize the discrete optimisation algorithm time shifts ($\mathcal{T}$). 
            \item[(ii)] Next, randomly select a small number of impulses $m$, with the value of $m$ treated as a hyperparameter. For each impulse, a random time shift is drawn from the $\mathcal{T}$ distribution creating a proposal vector. 
            \item [(iii)]The likelihood (both time shift and prediction error) for the proposal is calculated. 
            \item[(iv)] If the proposal likelihood is higher than the current maximum likelihood, our estimate is updated.
            \item [(v)] Return to (i) and repeat. The best estimate of the set of $\mathcal{T}$ improves each iteration. The number of iterations ($N$) is also treated as a hyperparameter.
        \end{enumerate} 
    \item [(3)] Optimise the model parameters $\theta_{f}$ and $\theta_{\mathcal{T}}$.
\end{enumerate}

@@

\subsection{Constraints}
Finally, a note on problem constraints. The likelihood estimates are dependent on the inner optimisation procedure. There is no guarantee that the global maximum will be found, particularly for sequences with a high density of impulses. 
When the density (in time) of non-zero impulses in $x(t)$ is high, the possible solution space of $\mathcal{T}$ grows exponentially.
In such cases, the estimated likelihood is likely to be close to, but not exactly the same as the real value. Therefore, TVS Regression is most appropriate for sparse time series inputs.

@@

The following section demonstrates a simulated univariate example of TVS Regression. Figure \ref{fig:simulated_example} shows the simulated time series. The full code can be found via \href{https://github.com/aaron1rcl/tvs_regression/blob/master/notebooks/1_univariate_example.ipynb}{TVS Regression on GitHub}. 
The input signal has $20$ non-zero impulses which have been drawn from the standard normal distribution. The system is 'off' when $x(t) =0$. In other words, when $x(t) = 0$, $f(x(t)) = 0$. The true shift distribution $\mathcal{T}$ is given by $\mathcal{T} \sim Poisson(\lambda = 2)$. There is one value of $\tau$ for each of the 20 impulses. The green line in Figure \ref{fig:simulated_example} represents the shifted series, corresponding to the time at which the effect occurs. The blue line is the output $y$ which includes a small amount of gaussian noise and an intercept. The values for the parameters were arbitrarily selected. The output $y$ is defined by the following equation:

@@

\begin{center}
\begin{table}
\begin{tabular}{|c|c|c|c|}
\hline
Parameter & TVS Regression & OLS Regression & True Value \\
\hline
$\beta$ & 2.09 & 0.50 & 2.00  \\
\hline
$Intercept$ & 6.53 & 6.62 & 6.50 \\
\hline
$\lambda_{\tau}$ & 1.54 & $N/A$ & 1.40 \\
\hline
$\sigma_{\epsilon}$ & 0.20 & 1.03 & 0.20 \\
\hline
\end{tabular}
\caption{ Parameter comparison between TVS and OLS regression against the true value.} 
\label{results_table}
\end{table}
\end{center}
